---
title: "Final Project"
author: "Estephani Barroso, Kassandra Chino Gonzalez, Yu-Fang Huang"
date: "2023-11-07"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Team members: 
Estephani Barroso, Kassandra Chino Gonzalez, Yu-Fang Huang

# Date set: 
COVID-19 Daily Counts of Cases

# url:
<https://data.cityofnewyork.us/api/views/rc75-m7u3/rows.csv?accessType=DOWNLOAD>

# Description of your data set: 
This data set counts the daily number of New York City residents who tested positive for COVID-19, were hospitalized with COVID-19, and the number of deaths among COVID-19 patients. This data set includes 67 variables and 1158 observations. These observations are recorded from 2/29/2020 to 5/1/2023. All of the variables are integer values. For the first few columns, it counts the daily number of cases and deaths in New York City, then counts the average seven days of those records. After that, it broke down into five different regions, which are Brooklyn (BK), Bronx (BX), Manhattan (MN), Queens (QU), and Staten Island (SI). Each region also counts the total cases and deaths and the average seven days of those cases.

# Questions: 
Which of the five regions has the highest death in covid 19?
On which day has the most COVID-19 cases?
Which month in each year had the most cases in NYC? Most deaths?
Visualization of NYC COVID-19 cases and deaths breakdown by year and then month in each year
What is the average number of deaths and COVID-19 cases on a yearly, monthly, weekly basis in NYC?


# Marginal summaries:

```{r}
library(readr)
library(tidyverse)

Data <- read.csv("COVID.csv")

head(Data)
dim(Data) # 1158 observations and 67 variables
str(Data) # All integer variables with data being the only character variable 

missing_values <- colSums(is.na(Data))
missing_values #no missing values 

Data$date_of_interest <- lubridate::mdy(Data$date_of_interest)


summary(Data)
Final <- Data %>%
  select(-contains("7DAY_AVG"))

```

```{r}
RegionsTotalDeaths <- Final %>% 
  summarize(SI_DEATH = sum(SI_DEATH_COUNT),
            QN_DEATH =sum(QN_DEATH_COUNT),
            MN_DEATH = sum(MN_DEATH_COUNT),
            BK_DEATH = sum(BK_DEATH_COUNT),
            BX_DEATH = sum(BX_DEATH_COUNT))
                              

Final2 <- Final %>%
  mutate(
    Year = year(date_of_interest),
    Month = month(date_of_interest, label = TRUE),
    Day = day(date_of_interest)
  )

CasesTotal <- Final2 %>% group_by (Month, Year) %>% summarize(TotalCases = sum(CASE_COUNT)) %>% 
  group_by(Year) %>% 
  slice(which.max(TotalCases))
                                                         
DeathsTotal <- Final2 %>% group_by (Month, Year) %>% summarize(TotalDeaths = sum(DEATH_COUNT))  %>% 
  group_by(Year) %>% 
  slice(which.max(TotalDeaths))  


ggplot(Final2, aes(x = factor(Year), y=DEATH_COUNT)) +
  geom_bar(stat = "identity", fill = "red", alpha = 0.7) +
  labs(title = "Deaths Per Yaer",
       x = "Year",
       y = "Deaths")

ggplot(Final2, aes(x = factor(Month), y=DEATH_COUNT)) +
  geom_bar(stat = "identity", fill = "red", alpha = 0.7) +
  labs(title = "Deaths Per Month",
       x = "Month",
       y = "Deaths")

ggplot(Final2, aes(x = factor(Month), y = DEATH_COUNT)) +
  geom_bar(stat = "identity", fill = "red", alpha = 0.7) +
  labs(title = "Deaths Per Month",
       x = "Month",
       y = "Deaths") +
  facet_grid(~Year, scales = "free_x", space = "free_x")

  
Final2$Year <- factor(Final2$Year, levels = c(2020, 2021, 2022, 2023))

                                                                          
```
